Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-07 inst/paper/main.tex
typo
Numbers and figures for usage metrics sections now in the .tex

We recorded \AllAllCallCountRnd calls to \eval and its variants (XXX001
for \eval, XXX002 for \c{evalq}, XXX003 for \c{eval.parent} and XXX004 for
\c{local}).

XXX005 packages in our corpus call \eval at least once.

\mypara{Number of eval call sites.}
We observed that the number of \eval call sites is small.
Figure XXX006 shows a violin plot of the distribution of call sites to
\eval and its three variants. For packages that have call sites, the
median is XXX007 call sites. \OneCallSitesEnrich{} packages have a
single call site, and one package, \emph{metafor}, has
\MaxCallSitesEnrich{} call sites. \NoCallSitesEnrich{} packages have
no calls to \eval. These numbers are under-approximations as \eval may
be aliased, called reflectively and given as argument to a
higher-order function such as \c{map}.

\mypara{Number of calls to eval.} We observed an average XXX010 calls
to \eval per run. \NOTE{this is in packages that had any call to eval.
  How do we compute this? Say we are interested in package P. Do we
  only count runs that call eval from package P. Or do we also count
  runs that did not have an eval but executed some code from P? The
  latter may be easier. Which one is right?} Figure XXX011 shows a
violin plot of the distribution of calls to eval per run. The largest
number of calls is XXX012 and comes from the package XXX013.

\mypara{Amount of code loaded by eval.} We use \c{deparse} to turn the
input to \eval into strings and measure the size of that input. The
largest input is XXX008 and the median input size is XXX009. Most
inputs are small, less than XXX014 bytes. The average size is XXX015
bytes.

\mypara{Amount of computation via eval.}
We compare the number of events intercepted during execution of \eval
across the whole corpus. For this, we use the number of interpreter
steps as a proxy for events. We observe \AllEventCountRnd events all
told, and \EvalEventCountRnd events occur inside eval
(\EvalEventAllPerc of all events).
Figure~XXX015 show a violin plot of events.

We observe a very wide spread in the number of events generated by
\eval calls. The largest number of events generated by an \eval call
is \EventsMaxCountRnd. However, a majority of \eval calls performs up
to \EventsMaxRangeA events suggesting that most expressions passed to
\eval are small. It is interesting to note that a higher proportion of
Core R \eval calls perform side effects in the
\EventsMinRangeB--\EventsMaxRangeB range. Only \EventsCoreEvalCountC
Core R \eval calls and \EventsPackageEvalCountC package \eval call
generate \EventsMinRangeC to \EventsMaxRangeC events. These \eval
calls originate from statistical modeling packages such as \mlogit,
\mboost, \metafor, \lavaan, \mclust and \gamlss.

@@

\myoara{Source.}
There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}

>>>>

We recorded \NbEvalAndVariantCallsRnd calls to \eval and its variants (\NbEvalCallsRnd
for \eval, \NbEvalqCallsRnd for \c{evalq}, \NbEvalParentCallsRnd for \c{eval.parent} and \NbLocalCallsRnd for
\c{local}).

\NbPackageEvalAtLeastOnce packages in our corpus call \eval at least once.

\mypara{Number of eval call sites.}
We observed that the number of \eval call sites is small.
Figure~\ref{fig:callSitesPerPackage} shows a violin plot of the distribution of call sites to
\eval and its three variants. For packages that have call sites, the
median is \NbMedianCallSites call sites. \NbOneCallSite{} packages have a
single call site, and one package, \emph{\MaxCallSitesPackage}, has
\MaxCallSites{} call sites. \PackagesNoCallsToEval{} packages have
no calls sites to \eval. These numbers are under-approximations as \eval may
be aliased, called reflectively and given as argument to a
higher-order function such as \c{map}.

\begin{figure}[h]
	\centering 
	\includegraphics[width=\textwidth]{call_sites_per_packages}
	\caption{Violin plot of the call sites per packages per eval function. } \label{fig:callSitesPerPackage}
\end{figure}

\mypara{Number of calls to eval.} We observed an average \AverageEvalCallsPerPackage{} calls
to \eval per run. \NOTE{this is in packages that had any call to eval.
  How do we compute this? Say we are interested in package P. Do we
  only count runs that call eval from package P. Or do we also count
  runs that did not have an eval but executed some code from P? The
  latter may be easier. Which one is right?} Figure~\ref{fig:evalCallsPerPackage} shows a
violin plot of the distribution of calls to eval per run. The largest
number of calls per run is \MaxEvalCalls{} and comes from the package \MaxEvalCallPackage.

\begin{figure}[h]
	\centering 
	\includegraphics[width=\textwidth]{eval_calls_per_packages}
	\caption{Violin plot of the normalized number of eval calls per package. } \label{fig:evalCallsPerPackage}
\end{figure}

\mypara{Amount of code loaded by eval.} We use \c{deparse} to turn the
input to \eval into strings and measure the size of that input. The
largest input is \MaxSizeEval and the median input size is \MedianSizeEval. Most
inputs are small, less than \SizeNineFiveRnd bytes. The average size is \AverageSizeEval
bytes.

\mypara{Amount of computation via eval.}
We compare the number of events intercepted during execution of \eval
across the whole corpus. For this, we use the number of interpreter
steps as a proxy for events. We observe \AllEventCountRnd{} \NOTE{Check that number.\quad} events all
told, and \NbEventsEvalRnd events occur inside eval
(\EvalEventAllPerc of all events).
Figure~\ref{fig:eventsPerPackage} show a violin plot of events.

\begin{figure}[h]
	\centering 
	\includegraphics[width=\textwidth]{nb_events_per_packages}
	\caption{Violin plot of the number of events per packages (log scale). } \label{fig:eventsPerPackage}
\end{figure}

We observe a very wide spread in the number of events generated by
\eval calls. The largest number of events generated by an \eval call
is \MaxiEventsEvalRnd. However, a majority (\SmallEventsEval) of \eval calls performs up
to 100 events suggesting that most expressions passed to
\eval are small. It is interesting to note that a higher proportion of
Core R \eval calls perform side effects in the
\EventsMinRangeB--\EventsMaxRangeB range. Only \EventsCoreEvalCountC
Core R \eval calls and \EventsPackageEvalCountC package \eval call
generate \EventsMinRangeC to \EventsMaxRangeC events. These \eval
calls originate from statistical modeling packages such as \mlogit,
\mboost, \metafor, \lavaan, \mclust and \gamlss.

@@

\mypara{Source.}
There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}
